master thesis

### 2018-03-21 Paper - Latex/Paper - Skizze - v4.tex
Paper - Results - Poisson Regression

%---------------------------------------------------------------------------------------------------
%---------------------------------------------------------------------------------------------------
%---------------------------------------------------------------------------------------------------
%---------------------------------------------------------------------------------------------------
\chapter{Results}
  A big point in this chapter is the result of examining the input data. How is the situation? What 
    influence does that have on our actual result? What can we do about it? Our simulation gave us 
    some results to this. \\
  Otherwise, this chapter is mainly about the actual reconstruction of the states. This means, on 
    one hand investigation of origins and losses of the inner nodes and on the other, the prediction 
    of unknown states of leaf nodes. \\
   
  %---------------------------------------------------------------------------------------------------
  %---------------------------------------------------------------------------------------------------
  %---------------------------------------------------------------------------------------------------
  \section{Metadata analysis}
    %---------------------------------------------------------------------------------------------------
    %---------------------------------------------------------------------------------------------------
    \subsection{Missing information}
      As previously presented, we have two types of missing information: unknown states of leaf nodes 
        and multifurcation. \\

      We examined the ridge of multifurcation of the tree. A complete phylogenetic tree would be 
        binary, which means the number of leaf nodes is closely to the number of internal nodes. But 
        since we only work with a synthesis tree, this tree is multifurcated: we have 241 974 internal 
        nodes and 2 293 463 leaf nodes. \\
      \begin{figure}[h]
        \centering
        \begin{subfigure}[b]{0.5\textwidth}
          \includegraphics[width=\textwidth]{Figures/multifurc.pdf}
          \caption{Histogram with automatic binwidth. \\ ~ \\ ~}
        \end{subfigure}
        ~~~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. 
          %(or a blank line to force the subfigure onto a new line)
        \begin{subfigure}[b]{0.42\textwidth}
          \includegraphics[trim = 0mm 0mm 30mm 0mm, clip, width=\textwidth]{Figures/multifurc_small.pdf}
          \caption{Histogram with $binwidth = 1$. \\ light blue: binary; dark blue: multifunction}
        \end{subfigure}
        \caption{Histograms about the multifurcation of the internal nodes of the synthesis tree.}
        \label{fig:childrenOfNodes}
      \end{figure}

      For a first overview we collected for every node its number of children (degree $-1$), and plotted
        this in two histograms, see figure \ref{fig:childrenOfNodes}. \\
      The multifurcation affects only the internal nodes. We collected the number of children $-2$ of 
        every node (a node with two children is binary). That means it discribes the number of nodes which we have lost from the real (binary) 
      phylogenetic tree. \\
      As \colorbox{red}{you} can see, we are very far from a binary tree. \\

      Some subtrees have been examined, these have the following percentages of missing information: See 
        table \ref{table:percentage loss information subtrees}.

      \begin{table} [h]
        \begin{center}
          \begin{tabular}{ |l|r|r||r|r|r| }
            \hline
            Domain  & \% unknown states & \% multifurcation \\ 
            / Kingdom  & (missing state information & (missing \\
            / Phylum / Class & of leaf nodes) & internal nodes) \\
            \hline \hline
            Eukaryota       & 97.34~\%  & 98.16~\% \\
            \hline \hline
            Metazoa         & 96.44~\%  & 87.93~\% \\ \hline
            Fungi           & 98.87~\%  & 96.97~\% \\ \hline
            Chloroplastida  & 99.14~\%  & 89.46~\% \\
            \hline \hline            
            % Arthropoda      & 97.49~\%  & 89.95~\% \\ \hline
            Apicomplexa     & 86.26~\%  & 87.16~\% \\ \hline
            Nematoda        & 89.01~\%  & 88.59~\% \\ \hline
            Chordata        & 88.59~\%  & \cellcolor{green!50}66.49~\% \\ \hline
            Platyhelminthes & \cellcolor{green!50}68.73~\%  & 80.34~\% \\
            \hline \hline            
            Insecta         & 97.11~\%  & 90.78~\% \\
            \hline  
          \end{tabular}
        \end{center}
        \caption{Examination of subtrees regarding missing information}
        \label{table:percentage loss information subtrees} 
      \end{table}

      %---------------------------------------------------------------------------------------------------
      \subsubsection{Data artifacts}
        At this point we also found out that there are some nodes with only one child node (55,700 nodes). \\
        These are both, the most nodes are right in front of a leaf, as well as some nodes are deep in the 
          tree (3,956 with height $>2$). They are probably a result from the fact that taxonomic information 
          has been incorporated into a phylogeny. \\
        Some examples:
        \begin{itemize}
          \item Nephroselmidophyceae: (class) \\
            https://tree.opentreeoflife.org/opentree/argus/ottol@1038762
          \item Phrynocrinidae: (family) \\
            https://tree.opentreeoflife.org/opentree/argus/ottol@3647979
          \item Elaeocarpus sylvestris: \\
            https://tree.opentreeoflife.org/opentree/argus/opentree9.1@ott166969
        \end{itemize}

      %---------------------------------------------------------------------------------------------------
      \subsubsection{Taxa}
        The investigation of the taxonomy revealed that our tree has three kingdoms: Chloroplastida, 
          Metazoa, Fungi, 53 phyla, 195 classes and 924 orders. \\
        Since the analysis of the tree is not part of this work, it should be mentioned here that, 
          according to recent findings, this is not complete and we lack some taxa in every rank. For 
          example, Cavalier-Smith says that one distinguishes between seven and nine kingdoms 
          \cite{CavalierSmith1981}. \\
        In section \pageref{subsec:listPhyla} of the appendix \colorbox{red}{you} can find a list of all phyla.

      %---------------------------------------------------------------------------------------------------
      \subsubsection{Poisson regression of the multifurcation}
        % glm (generalized linear model) analysis: \\
        The intercept is $2.821 > 0$ $\Rightarrow$ there is a multifunction.
        (Intercept: Stärke der Multifurcation) \\
        Comparing the different kingdoms, we find that multifunctionality is greater in Fungi than in 
        Chloroplastida than in Metazoa:
        $$4.0999 (Fungi Intercept) > -0.9132 (Chloroplastida Intercept) > -1.4320 (Metazoa Intercept)$$
        % Degrees of Freedom: 96010 Total (i.e. Null);  96010 Residual
        Wir haben außerdem drei komplexitätsstärken von Modellen verglichen bezüglich der höhe und tiefe 
          des Baums mit dem folgenden Deviance Table:

        \begin{table}[h]
          \begin{center}
              \begin{tabular}{ |l|r|r|r|r|r| }
                \hline
                Model / Taxa & Kingdom & Phylum & Class & Order & Family \\
                \hline \hline
                multifurc $\sim$ taxa & 7774454 & \cellcolor{green!10}7435700 & \cellcolor{green!15}7337241 & \cellcolor{green!30}7076068 \\
                \hline
                multifurc $\sim$ taxa + depth & 7752303 & \cellcolor{green!10}7431609 & \cellcolor{green!15}7334754 & \cellcolor{green!30}7027578 \\
                multifurc $\sim$ taxa + max.height & 7730196 & \cellcolor{green!15}7375889 & \cellcolor{green!20}7275856 & \cellcolor{green!30}7005424 \\
                multifurc $\sim$ taxa + min.height & \cellcolor{green!10}7472500 & \cellcolor{green!20}7233486 & \cellcolor{green!25}7144686 & \cellcolor{green!40}6890703 \\
                multifurc $\sim$ taxa + mean.height & \cellcolor{green!15}7304402 & \cellcolor{green!25}7128318 & \cellcolor{green!30}7055313 & \cellcolor{green!40}6815271 \\
                \hline
                multifurc $\sim$ taxa * depth & 7714881 & \cellcolor{green!15}7335396 & \cellcolor{green!20}7250759 & \cellcolor{green!40}6843004 \\
                multifurc $\sim$ taxa * max.height & \cellcolor{green!5}7692980 & \cellcolor{green!15}7311241 & \cellcolor{green!25}7187504 & \cellcolor{green!45}6795823 \\
                multifurc $\sim$ taxa * min.height & \cellcolor{green!10}7442387 & \cellcolor{green!25}7177002 & \cellcolor{green!30}7094933 & \cellcolor{green!45}6795099 \\
                multifurc $\sim$ taxa * mean.height & \cellcolor{green!20}7247309 & \cellcolor{green!30}7020258 & \cellcolor{green!35}6965794 & \cellcolor{green!50}6665565 \\
                \hline
              \end{tabular}
          \end{center}
          \caption{Residuals...}
          \label{table:...} 
        \end{table}
        * Residuals: Fehler - wieviele Werte sind nicht gut modelliert. (umso kleiner umso besser - grün) \\

        Interpretation: Die Multifurkation ist sehr ungleich verteilt. Daher ist die vorhersage umso 
          genauer umso kleinere Subtrees wir betrachten. ...

        Because of the difference in the complexity of the models, we compared their BICs:
        \begin{table}[h]
          \begin{center}
            \begin{tabular}{ |l|r|r|r|r|r| }
              \hline
              Model / Taxa & Kingdom & Phylum & Class & Order & Family \\
              \hline \hline
              multifurc $\sim$ taxa & 8273333 & \cellcolor{green!15}7937828 & \cellcolor{green!20}7842157 & \cellcolor{green!30}7644249 \\
              multifurc $\sim$ taxa & 8257680 & \cellcolor{green!15}7922207 & \cellcolor{green!20}7826490 & \cellcolor{green!35}7574154 \\
              \hline
              multifurc $\sim$ taxa + depth & 8273318 & \cellcolor{green!15}7934322 & \cellcolor{green!20}7839364 & \cellcolor{green!35}7539999 \\
              multifurc $\sim$ taxa + max.height & \cellcolor{green!15}7993515 & \cellcolor{green!25}7749121 & \cellcolor{green!30}7661817 & \cellcolor{green!40}7416211 \\
              multifurc $\sim$ taxa + min.height & 8251211 & \cellcolor{green!20}7875521  & \cellcolor{green!25}7778327 & \cellcolor{green!35}7516883 \\
              multifurc $\sim$ taxa + mean.height & \cellcolor{green!20}7825417 & \cellcolor{green!30}7644249 & \cellcolor{green!35}7572474 & \cellcolor{green!45}7340741 \\
              \hline
              multifurc $\sim$ taxa * depth & 8235932 & \cellcolor{green!20}7836755 & \cellcolor{green!25}7757688 & \cellcolor{green!45}7383808 \\
              multifurc $\sim$ taxa * max.height & \cellcolor{green!15}7963438 & \cellcolor{green!30}7693555 & \cellcolor{green!30}7614820 & \cellcolor{green!45}7335338 \\
              multifurc $\sim$ taxa * min.height & 8214030 & \cellcolor{green!20}7808940 & \cellcolor{green!30}7690618 & \cellcolor{green!45}7336627\\
              multifurc $\sim$ taxa * mean.height & \cellcolor{green!25}7768360 & \cellcolor{green!35}7536296 & \cellcolor{green!50}7484953 & \cellcolor{green!50}7206369 \\
              \hline
            \end{tabular} 
          \end{center}
          \caption{BIC...}
          \label{table:...} 
        \end{table}

      %---------------------------------------------------------------------------------------------------
      \subsubsection{Binomial regression of the unknown state information}

@@

    %---------------------------------------------------------------------------------------------------
    %---------------------------------------------------------------------------------------------------
    \subsection{Results of simulation / Influence of different parameters}

>>>>

%---------------------------------------------------------------------------------------------------
%---------------------------------------------------------------------------------------------------
%---------------------------------------------------------------------------------------------------
%---------------------------------------------------------------------------------------------------
\chapter{Results}
  A big point in this chapter is the result of examining the input data. How is the situation? What 
    influence does that have on our actual result? What can we do about it? Our simulation gave us 
    some results to this. \\
  Otherwise, this chapter is mainly about the actual reconstruction of the states. This means, on 
    one hand investigation of origins and losses of the inner nodes and on the other, the prediction 
    of unknown states of leaf nodes. \\
   
  %---------------------------------------------------------------------------------------------------
  %---------------------------------------------------------------------------------------------------
  %---------------------------------------------------------------------------------------------------
  \section{Metadata analysis - Missing information}
    As previously presented, we have two types of missing information: unknown states of leaf nodes 
      and multifurcation. \\

    We examined the ridge of multifurcation of the tree. A complete phylogenetic tree would be 
      binary, which means the number of leaf nodes is closely to the number of internal nodes. But 
      since we only work with a synthesis tree, this tree is multifurcated: we have 241 974 internal 
      nodes and 2 293 463 leaf nodes. \\
    For a first overview we collected for every node its number of children (degree $-1$), and plotted
      this in two histograms, see figure \ref{fig:childrenOfNodes}. \\
    The multifurcation affects only the internal nodes. We collected the number of children $-2$ of 
      every node (a node with two children is binary). That means it discribes the number of nodes which we have lost from the real (binary) 
    phylogenetic tree. \\
    \begin{figure}[h!]
      \centering
      \begin{subfigure}[b]{0.5\textwidth}
        \includegraphics[width=0.9\textwidth]{Figures/multifurc.pdf}
        \caption{Histogram with automatic binwidth. \\ ~ \\ ~}
      \end{subfigure}
      ~~~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. 
        %(or a blank line to force the subfigure onto a new line)
      \begin{subfigure}[b]{0.42\textwidth}
        \includegraphics[trim = 0mm 0mm 30mm 0mm, clip, width=0.9\textwidth]{Figures/multifurc_small.pdf}
        \caption{Histogram with $binwidth = 1$. \\ light blue: binary; \\ dark blue: multifunction}
      \end{subfigure}
      \caption{Histograms about the multifurcation of the internal nodes of the synthesis tree.}
      \label{fig:childrenOfNodes}
    \end{figure}
    As \colorbox{red}{you} can see, we are very far from a binary tree. \\
    Some subtrees have been examined, these have the following percentages of missing information: See 
      table \ref{table:percentage loss information subtrees}.
    \begin{table}[h!]
      \begin{center}
        \begin{tabular}{ |l|r|r||r|r|r| }
          \hline
          Subtree of & unknown states & multifurcation \\ 
          \hline \hline
          Eukaryota       & 97.34~\%  & 98.16~\% \\
          \hline \hline
          Metazoa         & 96.44~\%  & 87.93~\% \\ \hline
          Fungi           & 98.87~\%  & 96.97~\% \\ \hline
          Chloroplastida  & 99.14~\%  & 89.46~\% \\
          \hline \hline            
          % Arthropoda      & 97.49~\%  & 89.95~\% \\ \hline
          Apicomplexa     & 86.26~\%  & 87.16~\% \\ \hline
          Nematoda        & 89.01~\%  & 88.59~\% \\ \hline
          Chordata        & 88.59~\%  & \cellcolor{green!50}66.49~\% \\ \hline
          Platyhelminthes & \cellcolor{green!50}68.73~\%  & 80.34~\% \\
          \hline \hline            
          Insecta         & 97.11~\%  & 90.78~\% \\
          \hline  
        \end{tabular}
      \end{center}
      \caption{Examination of subtrees regarding missing information. \\ 
        % Taxa of subtrees: Domain, Kingdom, Phylum, Class \\ 
        Unknown states: missing state information of leaf nodes \\ 
        Multifurcation: missing internal nodes}
      \label{table:percentage loss information subtrees} 
    \end{table}

    %---------------------------------------------------------------------------------------------------
    %---------------------------------------------------------------------------------------------------
    \subsection{Data artifacts}
      At this point we also found out that there are some nodes with only one child node (55,700 nodes). \\
      These are both, the most nodes are right in front of a leaf, as well as some nodes are deep in the 
        tree (3,956 with height $>2$). They are probably a result from the fact that taxonomic information 
        has been incorporated into a phylogeny. \\
      Some examples:
      \begin{itemize}
        \item Nephroselmidophyceae: (class) \\
          https://tree.opentreeoflife.org/opentree/argus/ottol@1038762
        \item Phrynocrinidae: (family) \\
          https://tree.opentreeoflife.org/opentree/argus/ottol@3647979
        \item Elaeocarpus sylvestris: \\
          https://tree.opentreeoflife.org/opentree/argus/opentree9.1@ott166969
      \end{itemize}

    %---------------------------------------------------------------------------------------------------
    %---------------------------------------------------------------------------------------------------
    \subsection{Taxa}
      The investigation of the taxonomy revealed that our tree has three kingdoms: Chloroplastida, 
        Metazoa, Fungi, 53 phyla, 195 classes and 924 orders. \\
      Since the analysis of the tree is not part of this work, it should be mentioned here that, 
        according to recent findings, this is not complete and we lack some taxa in every rank. For 
        example, Cavalier-Smith says that one distinguishes between seven and nine kingdoms 
        \cite{CavalierSmith1981}. \\
      In section \pageref{subsec:listPhyla} of the appendix \colorbox{red}{you} can find a list of all phyla.

    %---------------------------------------------------------------------------------------------------
    %---------------------------------------------------------------------------------------------------
    \subsection{Poisson regression of the multifurcation}
      First of all we measured the strength of the multifurcation with the help of the intercept: 
        $2.821 > 0$ $\Rightarrow$ there is a multifunction. \\
      Comparing the different kingdoms, we find that multifunctionality is greater in Fungi than in 
        Chloroplastida than in Metazoa:
      $$4.0999 (Fungi Intercept) > -0.9132 (Chloroplastida Intercept) > -1.4320 (Metazoa Intercept)$$
      % Degrees of Freedom: 96010 Total (i.e. Null);  96010 Residual
      Then we have 9 times 4 models of different complexity levels.
      \begin{table}[h!]
        \begin{center}
            \begin{tabular}{ |l|r|r|r|r| }
              \hline
              Model / Taxa & Kingdom & Phylum & Class & Order \\
              \hline \hline
              multifurc $\sim$ taxa & 7774454 & \cellcolor{green!10}7435700 & \cellcolor{green!15}7337241 & \cellcolor{green!30}7076068 \\
              \hline
              multifurc $\sim$ taxa + depth & 7752303 & \cellcolor{green!10}7431609 & \cellcolor{green!15}7334754 & \cellcolor{green!30}7027578 \\
              multifurc $\sim$ taxa + max.height & 7730196 & \cellcolor{green!15}7375889 & \cellcolor{green!20}7275856 & \cellcolor{green!30}7005424 \\
              multifurc $\sim$ taxa + min.height & \cellcolor{green!10}7472500 & \cellcolor{green!20}7233486 & \cellcolor{green!25}7144686 & \cellcolor{green!40}6890703 \\
              multifurc $\sim$ taxa + mean.height & \cellcolor{green!15}7304402 & \cellcolor{green!25}7128318 & \cellcolor{green!30}7055313 & \cellcolor{green!40}6815271 \\
              \hline
              multifurc $\sim$ taxa * depth & 7714881 & \cellcolor{green!15}7335396 & \cellcolor{green!20}7250759 & \cellcolor{green!40}6843004 \\
              multifurc $\sim$ taxa * max.height & \cellcolor{green!5}7692980 & \cellcolor{green!15}7311241 & \cellcolor{green!25}7187504 & \cellcolor{green!45}6795823 \\
              multifurc $\sim$ taxa * min.height & \cellcolor{green!10}7442387 & \cellcolor{green!25}7177002 & \cellcolor{green!30}7094933 & \cellcolor{green!45}6795099 \\
              multifurc $\sim$ taxa * mean.height & \cellcolor{green!20}7247309 & \cellcolor{green!30}7020258 & \cellcolor{green!35}6965794 & \cellcolor{green!50}6665565 \\
              \hline
            \end{tabular}
        \end{center}
        \caption{Residuals of multifurcation models}
        \label{table:Residuals multifurcation} 
      \end{table}
      In Table \ref{table:Residuals multifurcation} is a deviance table, where the different residuals 
        are listed. Since only residuals are comparable in complexity, we have also calculated the BIC 
        values of the models in table \ref{table:BIC multifurcation}. \\
      \begin{table}[h]
        \begin{center}
          \begin{tabular}{ |l|r|r|r|r| }
            \hline
            Model / Taxa & Kingdom & Phylum & Class & Order \\
            \hline \hline
            multifurc $\sim$ taxa & 8273333 & \cellcolor{green!15}7937828 & \cellcolor{green!20}7842157 & \cellcolor{green!30}7644249 \\
            multifurc $\sim$ taxa & 8257680 & \cellcolor{green!15}7922207 & \cellcolor{green!20}7826490 & \cellcolor{green!35}7574154 \\
            \hline
            multifurc $\sim$ taxa + depth & 8273318 & \cellcolor{green!15}7934322 & \cellcolor{green!20}7839364 & \cellcolor{green!35}7539999 \\
            multifurc $\sim$ taxa + max.height & \cellcolor{green!15}7993515 & \cellcolor{green!25}7749121 & \cellcolor{green!30}7661817 & \cellcolor{green!40}7416211 \\
            multifurc $\sim$ taxa + min.height & 8251211 & \cellcolor{green!20}7875521  & \cellcolor{green!25}7778327 & \cellcolor{green!35}7516883 \\
            multifurc $\sim$ taxa + mean.height & \cellcolor{green!20}7825417 & \cellcolor{green!30}7644249 & \cellcolor{green!35}7572474 & \cellcolor{green!45}7340741 \\
            \hline
            multifurc $\sim$ taxa * depth & 8235932 & \cellcolor{green!20}7836755 & \cellcolor{green!25}7757688 & \cellcolor{green!45}7383808 \\
            multifurc $\sim$ taxa * max.height & \cellcolor{green!15}7963438 & \cellcolor{green!30}7693555 & \cellcolor{green!30}7614820 & \cellcolor{green!45}7335338 \\
            multifurc $\sim$ taxa * min.height & 8214030 & \cellcolor{green!20}7808940 & \cellcolor{green!30}7690618 & \cellcolor{green!45}7336627\\
            multifurc $\sim$ taxa * mean.height & \cellcolor{green!25}7768360 & \cellcolor{green!35}7536296 & \cellcolor{green!50}7484953 & \cellcolor{green!50}7206369 \\
            \hline
          \end{tabular} 
        \end{center}
        \caption{BIC of multifurcation models}
        \label{table:BIC multifurcation} 
      \end{table}

      % * Residuals: Fehler - wieviele Werte sind nicht gut modelliert. (umso kleiner umso besser - grün) \\
      The Residuals give us the error of the model. If the value is small, our data will be well modeled. \\
      Within the complexity classes it can be seen that the mean height gives the best additional factor. \\

      Despite higher complexity, the BIC values are getting smaller from model to model, meaning that 
        the finest model available here is also the best one of these. Smaller taxa than orders (eg 
        family) were computationally too expensive to calculate.

    %---------------------------------------------------------------------------------------------------
    %---------------------------------------------------------------------------------------------------
    \subsection{Binomial regression of the unknown state information}

@@

  %---------------------------------------------------------------------------------------------------
  %---------------------------------------------------------------------------------------------------
  %---------------------------------------------------------------------------------------------------
  \section{Results of simulation / Influence of different parameters}
